CF Registry Analysis

### 2023-11-30 Christine Chan/CF Registry/cf_registry_analysis.qmd
fixed really stupid error that was removing a bunch of NAs

```{r}
# Import data and codebook
annual <- read.csv("~/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_Annualized_Del1.csv",
  na.strings = c("", "N/A")
)
data_dict <- readxl::read_excel("~/Documents/Work/Christine Chan/CF registry data/Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
# Remove those with T1D or T2D
annual <- annual %>% filter(cfrd_status_annual != 4, cfrd_status_annual != 5)
annual$cfrd_status_annual <- factor(annual$cfrd_status_annual,
  levels = 1:3,
  labels = c("CFRD-", "CFRD-", "CFRD+")
)
# Get mutations
demogs <- read.csv("~/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_DemogCFDiag_Del1.csv")
demogs <- demogs %>% select(eDWID, Gender, MutClass)
annual <- left_join(annual, demogs, by = join_by(eDWID))
# Variable formatting
# If A_supplement_feeding is 0, then A_feeding2 is also 0 because they are not
# on any supplemental feeding
# Include all kinds of tube feeds
annual$A_feeding2[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding3[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding4[annual$A_supplement_feeding == 0] <- 0
tube_fields <- c("A_feeding2", "A_feeding3", "A_feeding4")
annual$tube_feeding <-
  apply(annual[, tube_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Liver disease
liver_fields <- c("A_hepatobiliary1_3", paste0("A_hepatobiliary2_", 1:4))
annual$liver_disease <-
  apply(annual[, liver_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Calculate percentiles
annual$bmi <- annual$A_weight / ((annual$A_height / 100)^2)
annual$age_perc <- ifelse(annual$Age_YrEnd < 20, annual$Age_YrEnd, 20)
annual$height_perc <- sds(annual$A_height,
  age = annual$age_perc,
  sex = annual$Gender, male = "M", female = "F",
  item = "height2_20", ref = cdc.ref, type = "perc"
) * 100
annual$bmi_perc <- sds(annual$bmi,
  age = annual$age_perc,
  sex = annual$Gender, male = "M", female = "F",
  item = "bmi", ref = cdc.ref, type = "perc"
) * 100
# Format columns
annual$MutClass <- factor(annual$MutClass,
  levels = c("1-3", "4-5", "Oth", "N/A"),
  labels = c("1-3", "4-5", "Oth", "Unknown")
)
annual$NumPulmExacerbation <- cut(annual$NumPulmExacerbation,
  labels = c("0", "1", "2", ">=3"),
  breaks = c(0, 1, 2, 3, Inf), right = F
)
annual$A_FEV1_pct_predicted <- cut(annual$A_FEV1_pct_predicted,
  breaks = c(0, 40, 80, 100, Inf), right = F
)
annual$A_FEV1_pct_predicted <- relevel(annual$A_FEV1_pct_predicted, ref = "[100,Inf)")
annual$height_perc <- cut(annual$height_perc, breaks = c(0, 10, 25, 50, Inf))
annual$height_perc <- relevel(annual$height_perc, ref = "(50,Inf]")
annual$bmi_perc <- cut(annual$bmi_perc, breaks = c(0, 10, 50, 85, Inf))
annual$bmi_perc <- relevel(annual$bmi_perc, ref = "(50,85]")
annual$age_group <- cut(annual$Age_YrEnd, breaks = c(10, 18, 30, Inf), right = F)
annual$age_group <- relevel(annual$age_group, ref = "[10,18)")
# Bugs
bugs <- c(
  "A_pseudomonasaeruginosa", "A_fungalyeast1",
  "A_pulmonarycomplications1", "A_mycocultureresults3"
)
# Yes/No variables
yn_vars <- c(
  "A_IsOnEnzymes", "A_corticosteroids1", "A_feeding2",
  "A_Vx770", "A_VX809comb", bugs, "tube_feeding", "liver_disease"
)
annual[, yn_vars] <- lapply(annual[, yn_vars], factor, levels = 0:1, labels = c("No", "Yes"))
annual$A_IsOnEnzymes <- relevel(annual$A_IsOnEnzymes, ref = "Yes")
```

@@

Based on CFFPR annualized data.

```{r}
# Find approximate birth date
annual$approx_birth_year = annual$ReviewYear - ceiling(annual$Age_YrEnd)
annual = annual %>% group_by(eDWID) %>% 
  mutate(approx_birth_year = round(mean(approx_birth_year)))
# Create survival dataset
s <- annual %>%
  rename(end = ReviewYear) %>% # ReviewYear is the end of the interval
  group_by(eDWID) %>%
  mutate(start = lag(end)) %>% # Get interval start
  select(eDWID, start,end, Age_YrEnd, approx_birth_year,cfrd_status_annual) # Sort columns
s$start[is.na(s$start)] = s$approx_birth_year[is.na(s$start)]
# Age group as time-varying covariate
s$age_group <- cut(s$Age_YrEnd,
  breaks = c(-Inf, 10, 19, 31, Inf), right = F,
  labels = c("< 10", "[10,19)", "[19,31)", ">= 31")
)
# Fill down diagnosis
s$cfrd_status_annual[s$cfrd_status_annual=="CFRD-"] = NA
s = s %>% group_by(eDWID) %>% fill(cfrd_status_annual,.direction = "down")
s$cfrd_status_annual[is.na(s$cfrd_status_annual)] = "CFRD-"
# Remove records where start == end
s = s[s$end > s$start,]
# Survival fit
surv_obj = Surv(time = as.numeric(s$start),time2 = as.numeric(s$end),
                event = as.numeric(s$cfrd_status_annual)-1)
fit <- survfit(surv_obj ~ age_group,id = eDWID,data = s)
ggsurvplot(fit,risk.table = T,cumevents = T,
  cumcensor = T,
           fun = "event",xlim = c(2003, 2018))
```

>>>>

```{r}
# Import data and codebook
annual <- read.csv("~/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_Annualized_Del1.csv",
  na.strings = c("", "N/A")
)
data_dict <- readxl::read_excel("~/Documents/Work/Christine Chan/CF registry data/Data_Raw/Copy of Codes for CFFPR_2018.xlsx")
data_dict$var <- paste(data_dict$field_name, data_dict$code_value, sep = "_")
# Remove those with T1D or T2D
annual <- annual %>% arrange(eDWID, ReviewYear)
annual = annual[!annual$cfrd_status_annual %in% c(4,5),]
annual$cfrd_status_annual <- factor(annual$cfrd_status_annual,
  levels = 1:3,
  labels = c("CFRD-", "CFRD-", "CFRD+")
)
# Get mutations
demogs <- read.csv("~/Documents/Work/Christine Chan/CF registry data/Data_Raw/CFF19_DemogCFDiag_Del1.csv")
demogs <- demogs %>% select(eDWID, Gender, MutClass)
annual <- left_join(annual, demogs, by = join_by(eDWID))
# Variable formatting
# If A_supplement_feeding is 0, then A_feeding2 is also 0 because they are not
# on any supplemental feeding
# Include all kinds of tube feeds
annual$A_feeding2[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding3[annual$A_supplement_feeding == 0] <- 0
annual$A_feeding4[annual$A_supplement_feeding == 0] <- 0
tube_fields <- c("A_feeding2", "A_feeding3", "A_feeding4")
annual$tube_feeding <-
  apply(annual[, tube_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Liver disease
liver_fields <- c("A_hepatobiliary1_3", paste0("A_hepatobiliary2_", 1:4))
annual$liver_disease <-
  apply(annual[, liver_fields], 1, function(r) {
    if (all(is.na(r))) {
      return(NA)
    } else {
      return(max(as.numeric(r), na.rm = T))
    }
  })
# Calculate percentiles
annual$bmi <- annual$A_weight / ((annual$A_height / 100)^2)
annual$age_perc <- ifelse(annual$Age_YrEnd < 20, annual$Age_YrEnd, 20)
annual$height_perc <- sds(annual$A_height,
  age = annual$age_perc,
  sex = annual$Gender, male = "M", female = "F",
  item = "height2_20", ref = cdc.ref, type = "perc"
) * 100
annual$bmi_perc <- sds(annual$bmi,
  age = annual$age_perc,
  sex = annual$Gender, male = "M", female = "F",
  item = "bmi", ref = cdc.ref, type = "perc"
) * 100
# Format columns
annual$MutClass <- factor(annual$MutClass,
  levels = c("1-3", "4-5", "Oth", "N/A"),
  labels = c("1-3", "4-5", "Oth", "Unknown")
)
annual$NumPulmExacerbation <- cut(annual$NumPulmExacerbation,
  labels = c("0", "1", "2", ">=3"),
  breaks = c(0, 1, 2, 3, Inf), right = F
)
annual$A_FEV1_pct_predicted <- cut(annual$A_FEV1_pct_predicted,
  breaks = c(0, 40, 80, 100, Inf), right = F
)
annual$A_FEV1_pct_predicted <- relevel(annual$A_FEV1_pct_predicted, ref = "[100,Inf)")
annual$height_perc <- cut(annual$height_perc, breaks = c(0, 10, 25, 50, Inf))
annual$height_perc <- relevel(annual$height_perc, ref = "(50,Inf]")
annual$bmi_perc <- cut(annual$bmi_perc, breaks = c(0, 10, 50, 85, Inf))
annual$bmi_perc <- relevel(annual$bmi_perc, ref = "(50,85]")
annual$age_group <- cut(annual$Age_YrEnd, breaks = c(-Inf,10, 18, 30, Inf), right = F)
annual$age_group <- relevel(annual$age_group, ref = "[10,18)")
# Bugs
bugs <- c(
  "A_pseudomonasaeruginosa", "A_fungalyeast1",
  "A_pulmonarycomplications1", "A_mycocultureresults3"
)
# Yes/No variables
yn_vars <- c(
  "A_IsOnEnzymes", "A_corticosteroids1", "A_feeding2",
  "A_Vx770", "A_VX809comb", bugs, "tube_feeding", "liver_disease"
)
annual[, yn_vars] <- lapply(annual[, yn_vars], factor, levels = 0:1, labels = c("No", "Yes"))
annual$A_IsOnEnzymes <- relevel(annual$A_IsOnEnzymes, ref = "Yes")
```

@@

## Overall

```{r}
# Find the year that each person was diagnosed, make wide
first_cfrd <- annual %>%
  filter(cfrd_status_annual == "CFRD+") %>%
  group_by(eDWID) %>%
  slice_head(n = 1) %>%
  select(eDWID, ReviewYear) %>%
  mutate(v = 1) %>%
  pivot_wider(
    id_cols = eDWID, names_from = ReviewYear, values_from = v,
    values_fill = 0
  )
# Make a dataframe for who was at risk each year
at_risk <- annual %>%
  select(eDWID, ReviewYear) %>%
  mutate(v = 1) %>%
  pivot_wider(
    id_cols = eDWID, names_from = ReviewYear, values_from = v,
    values_fill = 0
  )
# Add age information
ages <- annual %>%
  pivot_wider(
    id_cols = eDWID, names_from = ReviewYear, values_from = age_group
  )
```
